Incidence and period prevalence of confirmed COVID-19 cases in care home staff and residents: Cohort study of secondary use records in a large care home provider in England, Scotland and Northern Ireland (March-June 2020)

### 2020-06-19 covidWarehouse/manuscript/descriptives.Rmd
manuscript version 0.1

---
title: "Incidence and period prevalence of confirmed COVID-19 cases in care home staff and residents: Cohort study of secondary use records in a large care home provider in England, Scotland and Northern Ireland (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
date: "16/06/2020"
bibliography: references.bib
csl: sage-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  pdf_document:
    citation_package: natbib
    df_print: kable
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

@@

```{r descriptives}
homes_country <- select(reference_homes, home_code, postcode) %>% 
  left_join(transmute(reference_geography, postcode, 
                      country, region_england = region), by = "postcode") %>% 
  mutate(region_UK = if_else(country == "England", region_england, country)) %>% 
  select(-postcode)

incidents <- incidents %>% 
  dplyr::filter(dplyr::between(incident_date, date_min, date_max)) #%>% 
  # dplyr::filter(!resident_id %in% 
  #                 residents$resident_id[residents$funding_type %in% c("Young Chronic Sick/YPD", 
  #                                                                     "Younger Person")])

@@

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max)) %>% 
  filter(!funding_type %in% c("Young Chronic Sick/YPD", "Younger Person"))

new_cases <- filter(new_cases, between(date, date_min, date_max_tallies))

@@

desc_homes <- reference_homes %>% 
  left_join(transmute(reference_geography, postcode, 
                      country, region_england = region), by = "postcode") %>% 
  left_join(filter(beds, year=="2020"), by = "home_code") %>% 
  left_join(res_counts, by = "home_code") %>%
  mutate(Scope = paste("FSHCG", country)) %>% 
  group_by(Scope) %>% 
  summarise(`Total homes` = n_distinct(home_code),
            `Total beds` = sum(beds),
            `Total contract beds` = sum(contract_beds, na.rm = T),
            `Total residents (excl. contract beds)` = sum(unique_residents, na.rm = T))

@@

desc_overall <- list()
desc_overall$total_residents <- dplyr::n_distinct(desc_residents$resident_id)
#COUNTS
desc_overall$CT_total_symptomatic <- sum(new_cases$new_symptomatic_home, na.rm=T) + 
  sum(new_cases$new_symptomatic_hospital, na.rm=T)
desc_overall$CT_p_symptomatic <- desc_overall$CT_total_symptomatic/desc_overall$total_residents
desc_overall$CT_p_symptomatic_ME <- 1.96 * sqrt(
  (desc_overall$CT_p_symptomatic*(1-desc_overall$CT_p_symptomatic)/desc_overall$total_residents)
)
#INCIDENT REPORTS
desc_overall$unique_residents_incidents <- dplyr::n_distinct(incidents$resident_id)

desc_overall$DAT_total_symptomatic <- n_distinct(filter(incidents, !is.na(covid_first_symptomatic))$resident_id)
desc_overall$DAT_p_symptomatic <- desc_overall$DAT_total_symptomatic/desc_overall$total_residents
desc_overall$DAT_p_symptomatic_ME <- 1.96 * sqrt(
  desc_overall$DAT_p_symptomatic * (1 - desc_overall$DAT_p_symptomatic) / 
    desc_overall$total_residents
)
desc_overall$total_tests <- nrow(dplyr::filter(incidents, covid_tested == 1))
desc_overall$unique_residents_tested <- dplyr::filter(incidents, covid_tested == 1) %>% 
  dplyr::n_distinct(.$resident_id)
desc_overall$unique_residents_test_result <- dplyr::filter(incidents, !is.na(covid_test_result)) %>% 
  dplyr::n_distinct(.$resident_id)
desc_overall$unique_residents_test_result_positive <- dplyr::filter(incidents, covid_test_result == 1) %>% 
  dplyr::n_distinct(.$resident_id)

@@

## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))

tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Hazard ratios of becoming a confirmed case estimated from a Cox Proportional Hazards model (n=", 
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ")"))

@@

fig_KM_CT_residents_symptomatic <- fig_nums(
  name = "KM_CT_residents_symptomatic",
  caption = paste0("Kaplan-Meier estimator of the cumulative probability of residents becoming symptomatic based on care home manager daily counts")
)
fig_KM_confirmed_residents <- fig_nums(
  name = 'fig_KM_confirmed_residents',
  caption = paste0("Kaplan-Meier estimator of the cumulative probability of residents (n=", 
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ") becoming confirmed cases based on Datix incidents report"))
fig_KM_CT_confirmed_residents <- fig_nums(
  name = 'fig_KM_CT_confirmed_residents',
  caption = "Kaplan-Meier estimator of the cumulative probability of residents becoming confirmed cases based on care home manager daily counts")

@@

@@

**Background**

**Methods:**

**Findings:** A total of `r formatbm(n_distinct(desc_residents$resident_id))` residents were examined across England, Scotland and Northern Ireland, `r tbrounding(mean(residents$home_type=="Nursing")*100)` of whom were in nursing homes and 39·2% received dementia care (`r table_nums("tab_resident_descriptives", display = "cite")`).

**Interpretation**

@@

COVID-19 causes significant mortality in elderly and vulnerable people and spreads easily in care homes where one in seven individuals aged > 85 years live. However, there is no surveillance for infection in care homes, nor are there systems (or research studies) monitoring the impact of the pandemic on individuals or systems. Usual practices are disrupted during the pandemic, and care home staff are taking on new and unfamiliar roles, such as advanced care planning. Understanding the nature of these changes is critical to mitigate the impact of COVID-19 on residents, relatives and staff.

This study reports on data obtained from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. In 2020, FSHCG represented approximately 9% of all beds in England, Scotland and Northern Ireland (`r table_nums("tab_national_descriptives", display = "cite")`)). FSHCG residents accessed nose and throat swab antigen tests during hospitalisations in the first instance, but tests became available in care homes from XXXXX. From 11 May 2020, FSHCG home undertook testing every resident. ***(Haydn, more detail would be greatly appreciated)***

@@

These counts include events relating to occupants of local authority contract beds.

@@

- figure for the cumulative number/proportion of staff and residents infected over time... this is another area of significant interest so we can get an estimate (with caveats) of the proportion of staff and residents who remain susceptible to infection. -> KM by age or care home type or nation/region
- how many care homes have not had any cases reported in staff or residents during the pandemic because these homes will presumably be at greatest risk in a second wave.
- individual and care home characteristics associated with high number of cases. Nationally the only 2 factors that have been proposed so far are care home size and provision of nursing care. -> **regression model on RR of infection** number of previous cases (lag effect) is important here. So the time effect

@@

The study population consisted of `r formatbm(desc_overall$total_residents)` residents across England, Scotland and Northern Ireland, `r tbrounding(mean(residents$home_type=="Nursing")*100)`% of whom were in nursing homes and 39·2% received dementia care (`r table_nums("tab_resident_descriptives", display = "cite")`). The vast majority (85·4%) were permanent residents. A total of `r formatbm(nrow(incidents))` COVID-19 incidents were reported as of `r format(max(incidents$incident_date), "%A %e %b %Y")` for  `r formatbm(desc_overall$unique_residents_incidents)` unique residents.

Datix reports and manager counts respectively indicated that `r formatbm(desc_overall$DAT_total_symptomatic)` (`r tbrounding(desc_overall$DAT_p_symptomatic*100)`% [95% confidence interval: `r tbrounding((desc_overall$DAT_p_symptomatic - desc_overall$DAT_p_symptomatic_ME)*100)`; `r tbrounding((desc_overall$DAT_p_symptomatic + desc_overall$DAT_p_symptomatic_ME)*100)`]) and `r formatbm(desc_overall$CT_total_symptomatic)` (`r tbrounding(desc_overall$CT_p_symptomatic*100)`% [`r tbrounding((desc_overall$CT_p_symptomatic - desc_overall$CT_p_symptomatic_ME)*100)`; `r tbrounding((desc_overall$CT_p_symptomatic + desc_overall$CT_p_symptomatic_ME)*100)`]) became symptomatic at some point between `r format(date_min, "%e %B %Y")` and `r format(date_max, "%e %B %Y")`. Temporal trends broken down by region (`r fig_nums("fig_KM_CT_residents_symptomatic", display = "cite")`) indicate wide variation across the UK, with care homes in the South East and Northern Ireland reporting in excess of a third of residents with symptoms.

@@

`r fig_nums("fig_KM_CT_residents_symptomatic")`

```{r KM_CT_residents_symptomatic, fig.height = 5, fig.width = 10}

@@

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_ct_residents_symp) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cum probability confirmed case",
                     breaks = seq(0,1,.05)) +
  scale_color_discrete("Region") +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_ct_residents_symp$date)),
               labels = format.Date(
                 unique(get_monday_date(km_ct_residents_symp$date)),
                 "%d/%m/%y"
               )) +
  theme_bw()

@@

@@

The reports were returned by `r formatbm(dplyr::n_distinct(incidents$home_code))` care homes  totalling `r formatbm(sum(beds[beds$year=="2020" & beds$home_code %in% unique(incidents$home_code), "beds"]))` beds. 
Datix incidents accounted for `r formatbm(desc_overall$unique_residents_tested)` residents tested at least once. Test results were known for `r formatbm(desc_overall$unique_residents_test_result)` residents: `r formatbm(desc_overall$unique_residents_test_result_positive)` residents had a record of a positive test result. This data cannot be treated as representative, as engagement with FSHCG confirmed that test coming back negative were not commonly recorded on Datix. Care home manager daily counts of new confirmed cases added up to `r formatbm(sum(new_cases$new_confirmed_home, na.rm = T))` cases tested in homes and `r formatbm(sum(new_cases$new_confirmed_hospital, na.rm = T))` tested in hospitals. The regional breakdown is presented in

The temporal trend of confirmed cases obtained from either Datix reports (`r fig_nums("fig_KM_confirmed_residents", display = "cite")`) or manager counts (`r fig_nums("fig_KM_CT_confirmed_residents", display = "cite")`)

`r fig_nums("fig_KM_confirmed_residents")`

```{r km_confirmed, fig.cap=fig_KM_confirmed_residents}
km_confirmed <- tline %>% 
  left_join(select(reference_homes, home_code, postcode)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  filter(rday == 1 & susceptible_confirmed == 1) %>% 
  mutate(home_type = factor(home_type, levels = c("Residential", "Nursing"))) %>% 
  group_by(resident_id, dob, gender, region_UK,
           home_type, user_type, funding_type, admission_type) %>% 
  summarise(
    start = as.numeric(min(date) - date_min),
    end = max(date)
  ) %>% 
  left_join(select(timelines, resident_id, covid_first_positive)) %>% 
  mutate(event = if_else(
    is.na(covid_first_positive),
    0L,
    as.integer(covid_first_positive == end)
  )) %>% 
  mutate(
    stop = as.numeric(end +1 - date_min),
    ageg = cut_age_denary(compute_age(dob, date_min), age.min = 50, age.max = 100)
  )

survfit(Surv(time = start, 
                   time2 = stop,
                   event = event) ~ region_UK, data = km_confirmed) %>% 
  ggsurvplot(., conf.int = TRUE, 
           fun = "event",
           legend = c(0.16, 0.6),
           ggtheme = theme_bw(),
           # TODO: double check labels at the end
           break.x.by	= 7,
           break.y.by = .02,
           ylim = c(0, .17),
           xlab  = paste0("Days since ", format(date_min, "%d %B %Y")),
           ylab = "Cum probability of confirmed case",
           legend.labs = sort(unique(km_confirmed$region_UK)),
           conf.int.alpha = .1)

@@

`r fig_nums("fig_KM_confirmed_residents")`

```{r KM_CT_residents_infection, fig.height = 5, fig.width = 10}

@@

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_ct_residents_confirmed) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cum probability confirmed case",
                     breaks = seq(0,1,.02)) +
  scale_color_discrete("Region") +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_ct_residents_confirmed$date)),
               labels = format.Date(
                 unique(get_monday_date(km_ct_residents_confirmed$date)),
                 "%d/%m/%y"
               )) +
  theme_bw()

@@

>>>>

---
title: "Incidence and period prevalence of confirmed COVID-19 cases in care home staff and residents: Cohort study of secondary use records in a large care home provider in England, Scotland and Northern Ireland (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
date: "16/06/2020"
bibliography: references.bib
csl: sage-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  pdf_document:
    citation_package: natbib
    df_print: kable
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

```{r descriptives}
reference_homes <- reference_homes %>% 
  left_join(select(filter(beds, year=="2020"), -year)) %>% 
  mutate(staff_to_bed_ratio = staff_total/beds,
         carer_to_bed_ratio = (reference_homes$staff_care_assistants + reference_homes$staff_nurses)/beds,
         bed_tot_category = cut(.$beds, 
           breaks = seq(20,100,15), right = F, 
           paste0(seq(20, 80, 15), "-", seq(34, 94, 15), " beds")))

homes_country <- select(reference_homes, home_code, postcode) %>% 
  left_join(transmute(reference_geography, postcode, 
                      country, region_england = region), by = "postcode") %>% 
  mutate(region_UK = if_else(country == "England", region_england, country)) %>% 
  select(-postcode)

incidents <- incidents %>% 
  dplyr::filter(dplyr::between(incident_date, date_min, date_max)) #%>%

@@

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max))

new_cases <- filter(new_cases, between(date, date_min, date_max_tallies)) %>% 
    filter(home_code %in% unique(residents$home_code))

@@

desc_homes <- reference_homes %>% 
  filter(home_code %in% unique(residents$home_code)) %>% 
  left_join(transmute(reference_geography, country, postcode), by = "postcode") %>% 
  left_join(res_counts, by = "home_code") %>%
  mutate(Scope = paste("FSHCG", country)) %>% 
  group_by(Scope) %>% 
  summarise(., 
            `Total homes` = n_distinct(home_code),
            `Total beds` = sum(beds),
            `Total contract beds` = sum(contract_beds, na.rm = T),
            `Total residents (excl. contract beds)` = sum(unique_residents, na.rm = T))

@@

desc_overall <- list()
desc_overall$total_residents <- dplyr::n_distinct(desc_residents$resident_id)
desc_overall$total_staff <- reference_homes %>% filter(home_code %in% unique(residents$home_code)) %>% .$staff_total %>% sum()

#COUNTS
desc_overall$CT_total_symptomatic <- sum(new_cases$new_symptomatic_home, na.rm=T) + 
  sum(new_cases$new_symptomatic_hospital, na.rm=T)
desc_overall$CT_p_symptomatic <- desc_overall$CT_total_symptomatic/desc_overall$total_residents
desc_overall$CT_p_symptomatic_ME <- 1.96 * sqrt(
  (desc_overall$CT_p_symptomatic*(1-desc_overall$CT_p_symptomatic)/desc_overall$total_residents)
)
desc_overall$CT_total_confirmed <-  sum(new_cases$new_confirmed_home, na.rm = T) +
  sum(new_cases$new_confirmed_hospital, na.rm = T)
desc_overall$CT_p_confirmed <- desc_overall$CT_total_confirmed/desc_overall$total_residents
desc_overall$CT_p_confirmed_ME <- 1.96 * sqrt(
  desc_overall$CT_p_confirmed * (1-desc_overall$CT_p_confirmed) / 
    desc_overall$total_residents
)
#INCIDENT REPORTS
desc_overall$unique_residents_incidents <- dplyr::n_distinct(incidents$resident_id)

desc_overall$DAT_total_symptomatic <- n_distinct(filter(incidents, !is.na(covid_first_symptomatic))$resident_id)
desc_overall$DAT_p_symptomatic <- desc_overall$DAT_total_symptomatic/desc_overall$total_residents
desc_overall$DAT_p_symptomatic_ME <- 1.96 * sqrt(
  desc_overall$DAT_p_symptomatic * (1 - desc_overall$DAT_p_symptomatic) / 
    desc_overall$total_residents
)

home_total_confirmed_DAT <- filter(incidents, !is.na(covid_first_symptomatic)) %>% 
  group_by(home_code) %>% 
  summarise(DAT_confirmed = tidyr::replace_na(n_distinct(resident_id), 0))
home_total_confirmed_CT <- new_cases %>% group_by(home_code) %>% 
  summarise(CT_confirmed = sum(tidyr::replace_na(new_confirmed_home, 0) + 
                                 tidyr::replace_na(new_confirmed_hospital, 0)))

home_total_confirmed <- distinct(reference_homes, home_code, home_name_clickview, postcode) %>% 
  left_join(select(reference_geography, postcode, country, region_UK), by = "postcode") %>% 
  left_join(res_counts, by = "home_code") %>% 
  left_join(home_total_confirmed_CT) %>% 
  left_join(home_total_confirmed_DAT) %>% 
  mutate(CT_confirmed = tidyr::replace_na(CT_confirmed, 0),
        DAT_confirmed = tidyr::replace_na(DAT_confirmed, 0))

home_zero_confirmed <- home_total_confirmed %>% 
  group_by(region_UK) %>% 
  summarise(`Total homes` = n(),
            zero_datix = sum(DAT_confirmed == 0),
            pct_datix = sum(DAT_confirmed == 0)/n(),
            zero_CT = sum(CT_confirmed == 0) ,
            pct_CT =  sum(CT_confirmed == 0)/n()) %>% 
  arrange(-pct_datix)

desc_overall$total_homes <- n_distinct(home_total_confirmed$home_code)
desc_overall$CT_zero_case <- sum(home_total_confirmed$CT_confirmed == 0)
desc_overall$DAT_zero_case <- sum(home_total_confirmed$DAT_confirmed == 0)

desc_overall$total_tests <- nrow(dplyr::filter(incidents, covid_tested == 1))
desc_overall$unique_residents_tested <- dplyr::filter(incidents, covid_tested == 1) %>% 
  dplyr::n_distinct(.$resident_id)
desc_overall$unique_residents_test_result <- dplyr::filter(incidents, !is.na(covid_test_result)) %>% 
  dplyr::n_distinct(.$resident_id)
desc_overall$unique_residents_test_result_positive <- dplyr::filter(incidents, covid_test_result == 1) %>%   dplyr::n_distinct(.$resident_id)

@@

## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_zero_cases <- table_nums(
  name = "tab_zero_cases",
  caption = "Number and proportion of homes without confirmed cases by UK region"
)
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ")"))

@@

fig_KM_CT_residents_symptomatic <- fig_nums(
  name = "KM_CT_residents_symptomatic",
  caption = paste0("Kaplan-Meier estimator of the cumulative probability of residents becoming symptomatic based on care home manager daily counts")
)
fig_KM_residents_confirmed <- fig_nums(
  name = 'fig_KM_residents_confirmed',
  caption = paste0("Kaplan-Meier estimator of the cumulative probability of residents (n=",
  formatbm(dplyr::n_distinct(desc_residents$resident_id)),
  ") becoming confirmed cases based on Datix incidents report"))
fig_KM_CT_confirmed_residents <- fig_nums(
  name = 'fig_KM_CT_confirmed_residents',
  caption = "Kaplan-Meier estimator of the cumulative probability of residents becoming confirmed cases based on care home manager daily counts")

fig_KM_staff_symptomatic <- fig_nums(
  name = 'fig_KM_staff_symptomatic',
  caption = paste0("Kaplan-Meier estimator of the cumulative probability of care home staff (n=",   formatbm(desc_overall$total_staff),
  ") becoming symptomatic cases based on manager-reported daily counts"))
fig_KM_staff_confirmed <- fig_nums(
  name = 'fig_KM_staff_confirmed',
  caption = paste0("Kaplan-Meier estimator of the cumulative probability of care home staff (n=",   formatbm(desc_overall$total_staff),
  ") becoming confirmed cases based on manager-reported daily counts"))

@@

```{r period_prevalence}
cis_data <- list(
  n_positive = 66,
  n_negative = 60618
)

england_results <- list()
england_results$n_positive = sum(inner_join(new_cases, filter(homes_country, country == "England"))[c("new_confirmed_home", "new_confirmed_hospital")], na.rm = T)
england_results$n_negative = n_distinct(inner_join(residents, filter(homes_country, country == "England"))$resident_id) - england_results$n_positive

RR_england <- (
  england_results$n_positive / (england_results$n_positive + england_results$n_negative)
  ) / (
  cis_data$n_positive / (cis_data$n_positive + cis_data$n_negative)
)

RR_SE <- sqrt(
  1/england_results$n_positive  + 
    1/cis_data$n_positive +
    1/(england_results$n_positive + england_results$n_negative) +
    1/(cis_data$n_positive + cis_data$n_negative)
)
```

@@

**Background** COVID-19 causes significant mortality in elderly and vulnerable people and spreads easily in care homes. Scarce evidence is available on epidemiological trends in this setting. This study reports on data obtained from the Four Seasons Health Care Group (FSHCG), an independent provider of long-term and respite residential and nursing care representing 9% of all beds in England, Scotland and Northern Ireland.

**Methods:** Secondary use datasets were obtained from FSHCG. Individual-level records on residents and COVID-19 related events are investigated to estimate the incidence and period prevalence of symptoms, confirmed case, and death. Findings are compared with aggregate counts reported daily by care home managers to FSHCG as well as with findings from a national community survey.

**Findings:** Records of `r formatbm(n_distinct(desc_residents$resident_id))` residents present in FSHCG homes between `r format(date_min, "%e %B %Y")` and `r format(date_max, "%e %B %Y")` were analysed. The period prevalence of COVID-19 symptoms is comprised between`r tbrounding(desc_overall$DAT_p_symptomatic*100)`% (based on incidents reports) and `r tbrounding(desc_overall$CT_p_symptomatic*100)`% (based on manager-reported counts). Across the period, managers registered a total of `r formatbm(desc_overall$CT_total_confirmed)` (`r tbrounding(desc_overall$CT_p_confirmed*100)`% [`r tbrounding((desc_overall$CT_p_confirmed - desc_overall$CT_p_confirmed_ME)*100)`; `r tbrounding((desc_overall$CT_p_confirmed + desc_overall$CT_p_confirmed_ME)*100)`]) cases confirmed with a naropharyngeal RNA test. Datix incidents reported just `r formatbm(desc_overall$CT_total_confirmed)` confirmed cases. In England, `r england_results$n_positive` confirmed cases occurred out of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. Comparing with the community, the relative risk of confirmed cases in England was `r tbrounding(RR_england)` [`r tbrounding(exp(log(RR_england) - 1.96 * RR_SE)) `; `r tbrounding(exp(log(RR_england) + 1.96 * RR_SE))`]. The staff-to-bed ratio was the only factor to be independently associated with the rate of confirmed cases: an increase by 10 percentage point in the number of staff relative to the bed total was associated with a 13% reduction [9%, 18%] in the rate of confirmed cases.

**Interpretation** Despite uncertain case ascertainment in care homes, internal surveillance suggests a greater transmission than in the community. Staffing may impact care homes' response to outbreaks and in turn influence rates of confirmed cases. The absence of high-quality case ascertainment hinders local infection prevention initiatives which could be better supported by laboratory data surveillance.

@@

COVID-19 causes significant mortality in elderly and vulnerable people and spreads easily in care homes where one in seven individuals aged 85 years and over live. However, there is no surveillance for infection in care homes, nor are there systems (or research studies) monitoring the impact of the pandemic on individuals or systems. Usual practices are disrupted during the pandemic, and care home staff are taking on new and unfamiliar roles, such as advanced care planning. Understanding the nature of these changes is critical to mitigate the impact of COVID-19 on residents, relatives and staff.

This study reports on data obtained from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. In 2020, FSHCG represented approximately 9% of all beds in England, Scotland and Northern Ireland (`r table_nums("tab_national_descriptives", display = "cite")`)). FSHCG residents underwent reverse transcriptase polymerase chain reaction (RT-PCR) viral antigen diagnostics based on nose and throat swab antigen tests during hospitalisations in the first instance, but such tests later became available in care homes from ***XXXXX***. From 11 May 2020, FSHCG home undertook testing every resident. ***(Haydn, more detail would be greatly appreciated)***

@@

These counts include events relating to occupants of local authority contract beds as well as occupants receiving a different category of support (eg. young people) who were excluded from the study population.

@@

The incidence of confirmed cases in England is contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between 26 April and 13 June 2020 [@cisbulletin2020; @cisdata2020].

@@

The study population consisted of `r formatbm(desc_overall$total_residents)` residents across England, Scotland and Northern Ireland, `r tbrounding(mean(residents$home_type=="Nursing")*100)`% of whom were in nursing homes and 39·2% received dementia care (`r table_nums("tab_resident_descriptives", display = "cite")`). The vast majority (85·4%) were permanent residents. A total of `r formatbm(nrow(incidents))` COVID-19 incidents were reported on Datix as of `r format(max(incidents$incident_date), "%A %e %b %Y")` for  `r formatbm(desc_overall$unique_residents_incidents)` unique residents. The reports were returned by `r formatbm(dplyr::n_distinct(incidents$home_code))` care homes  totalling `r formatbm(sum(beds[beds$year=="2020" & beds$home_code %in% unique(incidents$home_code), "beds"]))` beds.

Datix reports and manager counts respectively indicated that `r formatbm(desc_overall$DAT_total_symptomatic)` (`r tbrounding(desc_overall$DAT_p_symptomatic*100)`% [95% confidence interval: `r tbrounding((desc_overall$DAT_p_symptomatic - desc_overall$DAT_p_symptomatic_ME)*100)`; `r tbrounding((desc_overall$DAT_p_symptomatic + desc_overall$DAT_p_symptomatic_ME)*100)`]) and `r formatbm(desc_overall$CT_total_symptomatic)` (`r tbrounding(desc_overall$CT_p_symptomatic*100)`% [`r tbrounding((desc_overall$CT_p_symptomatic - desc_overall$CT_p_symptomatic_ME)*100)`; `r tbrounding((desc_overall$CT_p_symptomatic + desc_overall$CT_p_symptomatic_ME)*100)`]) residents became symptomatic at some point between `r format(date_min, "%e %B %Y")` and `r format(date_max, "%e %B %Y")`. Temporal trends broken down by region (`r fig_nums("fig_KM_CT_residents_symptomatic", display = "cite")`) indicate wide variation across the UK, with care homes in the South East and Northern Ireland reporting in excess of a third of residents with symptoms.

@@

```{r KM_CT_residents_symptomatic, fig.height = 4, fig.width = 7}

@@

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_ct_residents_symp) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cum probability symptomatic case",
                     breaks = seq(0,1,.05)) +
  scale_color_discrete("Region") +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_ct_residents_symp$date)),
               labels = format.Date(
                 unique(get_monday_date(km_ct_residents_symp$date)),
                 "%d/%m/%y"
               )) +
  theme_bw()  +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))

@@

`r fig_nums("KM_CT_residents_symptomatic")`

@@

Datix incidents reports accounted for `r formatbm(desc_overall$unique_residents_tested)` residents tested at least once. Test results were known for `r formatbm(desc_overall$unique_residents_test_result)` of them (`r tbrounding(desc_overall$unique_residents_test_result/desc_overall$unique_residents_tested*100)`%), including `r formatbm(desc_overall$unique_residents_test_result_positive)` residents with a positive test result. This data should be treated as representative, as tests coming back negative are not commonly recorded on Datix. Care home manager daily counts of new confirmed cases added up to `r formatbm(sum(new_cases$new_confirmed_home, na.rm = T))` cases tested in homes and `r formatbm(sum(new_cases$new_confirmed_hospital, na.rm = T))` tested in hospitals. The temporal trend of confirmed cases broken down by region is presented using both Datix reports (`r fig_nums("fig_KM_residents_confirmed", display = "cite")`) and manager counts (`r fig_nums("fig_KM_CT_confirmed_residents", display = "cite")`).

For England only, a total of `r england_results$n_positive` confirmed cases occurred out of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. In comparison, the survey of English community households @cisbulletin2020 found a total of `r cis_data$n_positive` confirmed cases out of `r formatbm(cis_data$n_positive + cis_data$n_negative)` sampled individuals. This implies a relative risk of `r tbrounding(RR_england)` [`r tbrounding(exp(log(RR_england) - 1.96 * RR_SE)) `; `r tbrounding(exp(log(RR_england) + 1.96 * RR_SE))`].

```{r km_confirmed, fig.height=4,fig.width=8}
km_confirmed <- tline %>% 
  left_join(select(reference_homes, home_code, postcode, 
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  filter(rday == 1 & susceptible_confirmed == 1) %>% 
  mutate(home_type = factor(home_type, levels = c("Residential", "Nursing"))) %>% 
  group_by(resident_id, dob, gender, region_UK, home_code,
           home_type, user_type, funding_type, admission_type, 
           staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category) %>% 
  summarise(
    start = as.numeric(min(date) - date_min),
    end = max(date)
  ) %>% 
  left_join(select(timelines, resident_id, covid_first_positive)) %>% 
  mutate(event = if_else(
    is.na(covid_first_positive),
    0L,
    as.integer(covid_first_positive == end)
  )) %>% 
  mutate(
    stop = as.numeric(end +1 - date_min),
    ageg = cut_age_denary(compute_age(dob, date_min), age.min = 50, age.max = 100)
  )

survfit(Surv(time = start, 
                   time2 = stop,
                   event = event) ~ region_UK, data = km_confirmed) %>% 
  ggsurvplot(., conf.int = TRUE, 
           fun = "event",
           legend = "right",
           ggtheme = theme_bw(),
           # TODO: double check labels at the end
           break.x.by	= 7,
           break.y.by = .02,
           ylim = c(0, .17),
           xlab  = paste0("Days since ", format(date_min, "%d %B %Y")),
           ylab = "Cum probability of confirmed case",
           legend.labs = sort(unique(km_confirmed$region_UK)),
           conf.int.alpha = .1)

@@

`r fig_nums("fig_KM_residents_confirmed")`

```{r KM_CT_residents_infection, fig.height = 4, fig.width = 7}

@@

ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = region_UK), data = km_ct_residents_confirmed) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.05, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cum probability confirmed case",
                     breaks = seq(0,1,.02)) +
  scale_color_discrete("Region") +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_ct_residents_confirmed$date)),
               labels = format.Date(
                 unique(get_monday_date(km_ct_residents_confirmed$date)),
                 "%d/%m/%y"
               )) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))

@@

`r fig_nums("fig_KM_CT_confirmed_residents")`

Between `r desc_overall$DAT_zero_case` (`r tbrounding( sum(home_zero_confirmed$zero_datix)/ sum(home_zero_confirmed[["Total homes"]])*100)`%) and  `r desc_overall$CT_zero_case` (`r tbrounding(sum(home_zero_confirmed$zero_CT)/sum(home_zero_confirmed[["Total homes"]])*100)`%) care homes had not accounted any confirmed cases whether in the home or in hospital (`r table_nums("tab_zero_cases", display="cite")`). Yet the significant conflict between the two sources of data makes it difficult to ascertain which care homes had not had any confirmed cases without querying the national laboratory information systems.

`r table_nums("tab_zero_cases")`
